### 2024-07-30 paper.tex
find what nilpotent elements of slodowy slice are. define "n-dim-slodowy-slice springer fiber"

\subsection{Finding the previous Slodowy slices}
First we find \(\ker \ad_f\).
We have \((fX)_{ij} = i (n - i) A_{i + 1,j}\), and \((Xf)_{ij} = (j - 1) (n - (j - 1)) A_{i, j - 1}\).
So, for all \(i, j \in \{1, ..., n\}\), we have 
\[i(n - i)A_{i + 1,j} = (j - 1)(n - (j - 1))A_{i, j - 1}.\]
Taking \(j = 1\), we find that \(A_{i,1} = 0\) for \(i \geq 2\).
Then, taking \(j > 1\), we find that for \(i, j \in \{1, ..., n - 1\}\),
\[A_{i + 1, j + 1} = \frac{(j - 1)(n - (j - 1))}{i(n - i)} A_{ij}.\]
So, \(\ker\ad_f\) is the set of matrices which are upper triangular and satisfy the above condition.
\par To find the Slodowy slice associated to the previous \(\spl_2\)-triple \((E, H, F)\), we just need to find \(\ker\ad_F\).
We have 
\[[F, X] = \begin{pmatrix}
    0 & 0 \\
    f X_{21} & fX_{22}
\end{pmatrix} - \begin{pmatrix}
    0 & X_{12} f\\
    0 & X_{22} f
\end{pmatrix}.\]
Thus, \(X_{21}\) must be all zeroes except for the first row, and \(X_{12}\) must be all zeroes except for the last column, and \(X_{22} \in \ker\ad_f\).
There is no restriction on \(X_{11}\).
This describes \(\ker\ad_F\).

@@

>>>>

\subsection{Finding the previous Slodowy slices}
First we find \(\ker \ad_f\).
We have \((fX)_{ij} = i (n - i) A_{i + 1,j}\), and \((Xf)_{ij} = (j - 1) (n - (j - 1)) A_{i, j - 1}\).
So, for all \(i, j \in \{1, ..., n\}\), we have 
\[i(n - i)A_{i + 1,j} = (j - 1)(n - (j - 1))A_{i, j - 1}.\]
Taking \(j = 1\), we find that \(A_{i,1} = 0\) for \(i \geq 2\).
Then, taking \(j > 1\), we find that for \(i, j \in \{1, ..., n - 1\}\),
\[A_{i + 1, j + 1} = \frac{(j - 1)(n - (j - 1))}{i(n - i)} A_{ij}.\]
So, \(\ker\ad_f\) is the set of matrices which are upper triangular and satisfy the above condition.
\par To find the Slodowy slice associated to the previous \(\spl_2\)-triple \((E, H, F)\), we just need to find \(\ker\ad_F\).
We have 
\[[F, X] = \begin{pmatrix}
    0 & 0 \\
    f X_{21} & fX_{22}
\end{pmatrix} - \begin{pmatrix}
    0 & X_{12} f\\
    0 & X_{22} f
\end{pmatrix}.\]
Thus, \(X_{21}\) must be all zeroes except for the first row, and \(X_{12}\) must be all zeroes except for the last column, and \(X_{22} \in \ker\ad_f\).
There is no restriction on \(X_{11}\).
This describes \(\ker\ad_F\).
\par For \(X \in \mathcal{S}_{m,n}\), define \(u(X) := X_{11}\).

\subsection{Finding \(\widetilde{\mathcal{N}}_{m,n}\)}
Let \(\mathcal{N}_m \subseteq \gl_m\) be the nilpotent elements.
Let \(\mathcal{S}_{m,n}'\) be the set of \(X \in \mathcal{S}_{m,n}\) such that both \(X\) and \(u(X)\) are nilpotent.
Let \(\widetilde{\mathcal{N}}_{m,n} = \{(\mathfrak{b}, X) : X \in \mathfrak{b}\} \subseteq \mathcal{B}_{m + n} \times \mathcal{S}_{m,n}'\).
Define \(\pi_{m,n} : \widetilde{\mathcal{N}}_{m,n} \to \mathcal{N}_m\) by \((\mathfrak{b}, X) \mapsto X_{11}\).
For \(Y \in \gl_m\), we call \(\pi_{m,n}^{-1}(Y)\) the \emph{n-dim-Slodowy-slice Springer fiber} at \(Y\).

@@

\begin{lemma}
    Let \(X \in \gl_m\), and let 
    \[Y = \begin{pmatrix}
        y_{11} & y_{12} & y_{13} & \cdots & y_{1,n-1} & y_{1n} \\
        d_1 & y_{22} & y_{23} & \cdots & y_{2,n-1} & y_{2n} \\
            & d_2   & y_{33} & \cdots & y_{3,n-1} & y_{3n} \\
            & & \ddots & \cdots & \vdots  & \vdots \\
            & & & d_{n - 2} & y_{n-1,n-1} & y_{n - 1,n}\\
            & & & & d_{n - 1} & y_{nn}
    \end{pmatrix} \in \gl_n.\]
    For any \(a, b \in \mathbb{C}^m\),
    \[\det 
    \begin{pNiceArray}{ccc|ccc}
        & & & & & \vert \\
        & X & & & & b    \\
        & & & & & \vert \\
       \hline
       \text{---} & a & \text{---} &  \\
       & & & & Y & \\
       & & &  \\
       \end{pNiceArray} = \]
       \[\det X \det Y + \left(\prod_i d_i\right) \det \begin{pNiceArray}{ccc|c}
            & & & \vert \\
            & X & & b     \\
            & & & \vert \\
            \hline 
        \text{---} & a & \text{---} & 0
       \end{pNiceArray}\]
\end{lemma}
\begin{proof}
    By induction on \(n\).
    In the case \(n = 1\), expanding along the last row (taking the usual interpretation of the empty product) gives the desired result.
    \par Now suppose \(n > 1\).
    Expanding along the last row, we get 
    \[d_{n - 1} \det \begin{pNiceArray}{ccc|ccc}
        & & & & & \vert \\
        & X & & & & b    \\
        & & & & & \vert \\
       \hline
       \text{---} & a & \text{---} &  \\
       & & & & Y_{n,n-1} & \\
       & & &  \\
       \end{pNiceArray} -
    y_{nn} \det \begin{pNiceArray}{ccc|ccc}
        & & & & & \\
        & X & & & &   \\
        & & & & & \\
       \hline
       \text{---} & a & \text{---} &  \\
       & & & & Y_{n,n} & \\
       & & &  \\
       \end{pNiceArray}.\]
    Using our inductive hypothesis for the first determiniant, and using that \(\det \begin{pNiceArray}{c|c}
        A_{11} & 0 \\
        A_{21} & A_{22}
    \end{pNiceArray} = \det A_{11} \det A_{22}\) for the second, the expression becomes
    \[d_{n - 1} \left(\det X \det Y_{n,n-1} + \left(\prod_{i \leq n - 2} d_i\right) \det \begin{pNiceArray}{ccc|c}
        & & & \vert \\
        & X & & b     \\
        & & & \vert \\
        \hline 
    \text{---} & a & \text{---} & 0
   \end{pNiceArray}\right)
   - y_{nn} \det X \det Y_{nn} = \]
   \[(d_{n - 1} Y_{n,n-1} - y_{nn} \det Y_{nn})\det X + \left(\prod_i d_i\right) \det \begin{pNiceArray}{ccc|c}
        & & & \vert \\
        & X & & b     \\
        & & & \vert \\
        \hline 
    \text{---} & a & \text{---} & 0
   \end{pNiceArray} = \]
   \[\det Y \det X + \left(\prod_i d_i\right) \det \begin{pNiceArray}{ccc|c}
        & & & \vert \\
        & X & & b     \\
        & & & \vert \\
        \hline 
    \text{---} & a & \text{---} & 0
   \end{pNiceArray}.\]
\end{proof}

\begin{corollary}
    If \(X\) is nilpotent, and
    \[\begin{pNiceArray}{ccc|ccc}
        & & & & & \vert \\
        & X & & & & b    \\
        & & & & & \vert \\
       \hline
       \text{---} & a & \text{---} &  \\
       & & & & Y & \\
       & & &  \\
       \end{pNiceArray}\]
    is nilpotent  as well, then \(Y\) is nilpotent (TODO: and that other determinant is zero).
    \begin{proof}
        By the previous lemma, the characteristic polynomial of the big matrix is
        \[g_X(\lambda) g_Y(\lambda) + f(\lambda),\]
        where \(g_X(\lambda) = \lambda^m\) is the characteristic polynomial of \(X\), and \(f(\lambda)\) is some polynomial of degree at most \(m - 1\).
    \end{proof}
\end{corollary}
\par Now, taking the previous corolloary and the first lemma together, we see that 
\[\mathcal{S}_{m,n}' = \left\{ \begin{pNiceArray}{ccc|cccc}
    & & & & & & \vert \\
    & X & & & & & b    \\
    & & & & & & \vert \\
   \hline
   \text{---} & a & \text{---} & 0 \\
    & & & 1 & 0  & \\
    &  & & & \ddots & \ddots  \\
    & & & & & 1 & 0
   \end{pNiceArray} : a, b \in \mathbb{C}^m, X \in \gl_m \textrm{ is nilpotent}\right\}.\]
